\LARGE \bf On the use of vehicle categorization and context information for messages forwarding in VANETs

### 2017-01-09 paperITSC.tex
Kifayat's changes João's changes

\author{João B. Ribeiro$^{1}$, Wellington D. Branquinho Junior$^{1}$, Kifayat Ullah$^{1}$ and Edson dos Santos Moreira$^{1}$
%\thanks{*This work was supported by \textcolor{green}{National Council for the Improvement of Higher Education} (CAPES)}
\thanks{$^{1}$João B. Ribeiro, Wellington D. Branquinho Junior, Kifayat Ullah and Edson dos Santos Moreira are with Instituto de Ciências Matemáticas e de Computação (ICMC), University of São Paulo (USP), São Carlos, Brazil
    \tt\small \{joao.b, wellingtonjr\}@usp.br and \{kifayat, edson\}@icmc.usp.br}
}

@@

With the increasing number of vehicles on the roads, traffic related problems like congestion, traffic jams, environmental pollution, and accidents, are very common these days. To solve transportation-related problems and make the driving experience safer, comfortable and enjoyable, VANETs were developed. A large number of applications ranges from safety to non-safety purposes are envisioned for VANETs. These applications make use of opportunistic encounters to forward messages. However, the unique characteristics of VANETs, make the opportunistic forwarding a challenging task. In this paper, we proposed a novel protocol called Message Forwarding by Categorization of Vehicles (\emph{MFCV}). Our protocol exploits contextual information and vehicles categorization to improve opportunistic forwarding of messages. We validate and compare our protocol with the Epidemic routing protocol, by means of a series of simulation experiments. Our protocol performed very close to the Epidemic protocol, in terms of messages delivery. Additionally, our results showed improved performance when vehicles categorization was considered in the forwarding decision. Finally, (\emph{MFCV}) performed better in terms of message latency, and resource utilization.

@@

Vehicular Ad hoc NETworks (VANETs) are a subclass of Mobile Ad hoc NETworks with unique characteristics like high speed of nodes and dynamic network topology. To disseminate data, VANETs enable communications between mobile nodes and infrastructure network. These networks support a large number of applications such as accident avoidance, intelligent traffic control, and data dissemination, thereby, providing safety, comfort, and convenience to the users \cite{conprova2013}.

@@

Motivated by the works presented in \cite{Yokoyama_2014, ullahadvertising} that make use of beacon messages and opportunistic vehicles communications to exchange information, we propose Message Forwarding by Categorization of Vehicles (\emph{MFCV}). \emph{MFCV} is an opportunistic message forwarding protocol that makes use of beacons and exploits contextual information, especially the vehicle usage category, to improve the messages delivery by consuming relatively low resources.

@@

In \cite{magf2009}, the authors addressed the problem of rapid topology changes and frequent disconnections in VANETs. They assumed that the vehicle is aware of its own movement and its neighbors, and uses this information to select the best next hop to carry the message. They concluded that using a combination of position, speed and direction have a significant performance improvement in terms of data delivery ratio and latency.

@@

By utilization the embedded sensors, a smart vehicle can produce rich context information. The context information about vehicle contains speed, direction, wheel friction, acceleration and position of the vehicle. Furthermore, a smart car would monitor its environment. The context information about the environment can be divided into static (e.g., road signs, number of lanes), and dynamic (e.g., wind speed, range of visibility, chances of rain). Using context information of the vehicle or environment make it possible to detect or prevent hazardous situations. Moreover, the context can also be useful in the knowledge distribution over a vehicular network \cite{tocadas2010}.

@@

For our work, we used the ontology\footnote{Draft available on (http://auto.sdo-gao.appspot.com/) Access in: 17 October 2016} proposed in \cite{gao_2016} that categorizes the vehicle according to its physical features (e.g., size of the vehicle, the number of axles), generating the \emph{physical category}. Furthermore, we also need to categorize the vehicles by its usage (i.e., private car, taxi, bus). Therefore, we have added in the ``\emph{CarUsageType}'' a vehicle \emph{usage category}.
This category defines how the vehicle will be used and allows the extraction of mobility patterns information. The \emph{physical category} should be used by traffic management companies, for instance, in the toll collection. While, the \emph{usage category} can be used by forwarding protocols to decide which node is the best to carry a message. In our protocol, the vehicle \emph{usage category} information is sent by vehicles, along with other information collected from sensors.

@@

In \emph{MFCV} we assume that each vehicle is able to collect the required contextual information about itself and store information received from its neighbors. Furthermore, the contextual information shared among the vehicles is reliable and the vehicles collaborate with the forwarding process since security is not the focus of this work. In the following subsections, we define the models used for the implementation.

@@

For network and transport layers, we used WAVE Short Message Protocol (WSMP) standard. We have adopted the IEEE 802.11p standard protocol for the PHY and MAC layers. We used the IEEE 1609.4 standard protocol for channel operation. This standard define two types of channel, the Control Channel (CCH) and the Service Channel (SCH). In your simulation all communication is through the service channel SCH1 174. The signal propagation was made according to the Simplified Path-Loss model, described in \cite{tse2005}.

@@

We developed a mobility scenario that consists of a grid of 9 km$^2$ (urban area) using the \emph{netgenerate} tool from SUMO. In order to evaluate the influence of the different categories of vehicles in the forwarding decision, 10 vehicles were classified as taxis and the remaining as private vehicles. We used \emph{randomTrips}, SUMO for generating routes, in which the taxis accomplished 15 km while the private vehicles only 2 km. Due to this, the taxi will take maximum simulation time (i.e., 10 minutes) to complete its route, whereas the private vehicles leave the scenario in approximately 2 minutes. This way, the taxis can increase the successful delivery rate of messages as compared to private vehicles.

@@

In the first experiment, we compared \emph{MFCV}, exploiting different context information, with the \emph{Epidemic} protocol measuring the number of messages that reached the destination. Since \emph{Epidemic} is a sort of controlled flooding, its successful delivery rate is very high, so our goal was to reach closer to its results.

In this experiment, a vehicle was chosen to create a message every 15 seconds, for delivery, generating a total of 30 messages. 
As seen in \autoref{grafico1} the combination \emph{DSC} (distance, speed, and category) shows an improvement of approximately 6\% when compared to the basic D (distance) and is 9\% below the performance of \emph{Epidemic}. Increasing the number of generated message to 60 with the TTL equal to 60 seconds, we get same results for \emph{D}, \emph{DS} and \emph{DSC}, but 6\% below the \emph{Epidemic} performance. Increasing the TTL to 120 seconds, the results of \emph{DSC} shows 5\% better performance than \emph{D} and \emph{DS}, but 12\% below \emph{Epidemic}. We compared both \emph{Epidemic} and \emph{DSC}, with 60 generated messages, increasing message TTL to 150 seconds, and the results displayed in \autoref{grafico2} show 100\% delivery success rate to \emph{Epidemic} and 98\% to our approach.

@@

This paper presented \emph{MFCV} - a message forwarding protocol for VANETs. \emph{MFCV} exploits context information to improve the opportunistic forwarding of messages. The vehicle distance to the destination, its speed and categories (type and use), were used in the decision make to enhance the process. Vehicle usage category was used to extract mobility habits of each vehicle. We used path history of all vehicles to observe if they are approaching destination and \emph{RateTimeToSend} mechanism to control the messages forwarding rate and the number of hops. We compared our results with \emph{Epidemic} routing protocol, in terms of successful delivery rate, average latency and the number of message packets send per minute. \emph{MFCV} performed only 6 to 9\% worse than \emph{Epidemic} in terms of message delivery success. Additionally, in terms of average latency, our protocol performed 10 to 15\% better than \emph{Epidemic}. Finally, in terms of message packets sent, our protocol sent around 70\% fewer message packets, as compared with \emph{Epidemic}, in the worst case. Our results conclude that our protocol performed as good as \emph{Epidemic}. Moreover, it consumed fewer resources as well.

As a future work, we intend to evaluate the impact of buffer size, the number of messages generated and the number of vehicles using real-world mobility traces. in our forwarding scheme. Additionally, we intend to investigate other vehicles categories such as buses and trucks in different environments like city and highways. We also plan to exploit other context information such as acceleration and heading of vehicles to increase our protocol success rate in terms of message delivery and latency.

>>>>

\author{João B. Ribeiro$^{1}$, Wellington D. Branquinho Junior$^{1}$, Kifayat Ullah$^{2}$ and Edson dos Santos Moreira$^{1}$
%\thanks{*This work was supported by \textcolor{green}{National Council for the Improvement of Higher Education} (CAPES)}
\thanks{$^{1}$João B. Ribeiro, Wellington D. Branquinho Junior and Edson dos Santos Moreira are with Instituto de Ciências Matemáticas e de Computação (ICMC), University of São Paulo (USP), São Carlos, Brazil.$^{2}$Kifayat Ullah is with CECOS University, Peshawar, Pakistan.
    \tt\small \{joao.b, wellingtonjr\}@usp.br and \{kifayat, edson\}@icmc.usp.br}
}

@@

With the increasing number of vehicles on the roads, traffic related problems like congestion, traffic jams, environmental pollution, and accidents, are very common these days. To solve transportation-related problems and make the driving experience safer, comfortable and enjoyable, VANETs were developed. A large number of applications ranges from safety to non-safety purposes are envisioned for VANETs. These applications make use of opportunistic encounters to forward messages. However, the unique characteristics of VANETs, make the opportunistic forwarding a challenging task. In this paper, we proposed a novel protocol called Message Forwarding by Categorization of Vehicles (\emph{MFCV}). Our protocol exploits contextual information and vehicles categorization to improve opportunistic forwarding of messages. We validate and compare our protocol with the Epidemic routing protocol, by means of a series of simulation experiments. Our protocol performed very close to the Epidemic protocol, in terms of messages delivery. Additionally, our results showed improved performance when vehicles categorization was considered in the forwarding decision. Finally, \emph{MFCV} performed better in terms of message latency, and resource utilization.

@@

Vehicular Ad hoc NETworks (VANETs) are a subclass of Mobile Ad hoc NETworks with unique characteristics like high speed of nodes and dynamic network topology. To disseminate data, VANETs enable communications between mobile nodes and infrastructure network. These networks support a large number of applications such as accident avoidance, intelligent traffic control, and data dissemination, thereby, providing safety, comfort, and convenience to the drivers \cite{conprova2013}.

@@

Motivated by the works presented in \cite{Yokoyama_2014, ullahadvertising, Ullah2016} that make use of beacon messages and opportunistic vehicles communications to exchange information, we propose Message Forwarding by Categorization of Vehicles (\emph{MFCV}). \emph{MFCV} is an opportunistic message forwarding protocol that makes use of beacons and exploits contextual information, especially the vehicle usage category, to improve the messages delivery by consuming relatively low resources.

@@

In \cite{magf2009}, the authors addressed the problem of rapid topology changes and frequent disconnections in VANETs. They assumed that the vehicle is aware of its own movement and its neighbors, and uses this information to select the best next hop to carry the message. They concluded that using a combination of vehicle position, speed and direction have a significant performance improvement in terms of data delivery ratio and latency.

@@

By utilizing the embedded sensors, a smart vehicle can produce rich context information. The context information about vehicle contains speed, direction, wheel friction, acceleration and position of the vehicle. Furthermore, a smart vehicle could monitor its environment. The context information about the environment can be divided into static (e.g., road signs, number of lanes), and dynamic (e.g., wind speed, range of visibility, chances of rain). Using context information of the vehicle or environment make it possible to detect or prevent hazardous situations. Moreover, the context can also be useful in the knowledge distribution over a vehicular network \cite{tocadas2010}.

@@

For our work, we used the ontology\footnote{Draft available on (http://auto.sdo-gao.appspot.com/) Access in: 17 October 2016} proposed in \cite{gao_2016} that categorizes the vehicle according to its physical features (e.g., size of the vehicle, the number of axles), generating the \emph{physical category}. Furthermore, we also need to categorize the vehicles by its usage (i.e., private car, taxi, bus). Therefore, we have added in the ``\emph{CarUsageType}'' a vehicle \emph{usage category}.
This category defines how the vehicle will be used and allows the extraction of mobility patterns information. The \emph{physical category} should be used by traffic management companies, for instance, in the toll collection. While, the \emph{usage category} can be used by forwarding protocols to decide which node is the best to carry a message. In our protocol, the vehicle \emph{usage category} information is send by vehicles, along with other information collected from sensors.

@@

In \emph{MFCV}, we assume that each vehicle is able to collect the required contextual information about itself and store information received from its neighbors. Furthermore, the contextual information shared among the vehicles is reliable and the vehicles collaborate in the forwarding process since security is not the focus of this work. In the following subsections, we define the models used for the implementation.

@@

For network and transport layers, we used WAVE Short Message Protocol (WSMP) standard. We have adopted the IEEE 802.11p standard protocol for the PHY and MAC layers. We used the IEEE 1609.4 standard protocol for channel operation. This standard define two types of channel, the Control Channel (CCH) and the Service Channel (SCH). In our simulation all communication is through the service channel SCH1 174. The signal propagation was made according to the Simplified Path-Loss model, described in \cite{tse2005}.

@@

We developed a mobility scenario that consists of a grid of 9 km$^2$ (urban area) using the \emph{netgenerate} tool from SUMO. In order to evaluate the influence of the different categories of vehicles in the forwarding decision, 10 vehicles were classified as taxis and the remaining as private vehicles. We used SUMO tool \emph{randomTrips} for generating routes, in which the taxis accomplished 15 km while the private vehicles only 2 km. Due to this, the taxi will take maximum simulation time (i.e., 10 minutes) to complete its route, whereas the private vehicles leave the scenario in approximately 2 minutes. This way, the taxis can increase the successful delivery rate of messages as compared to private vehicles.

@@

In the first experiment, we compared \emph{MFCV} by exploiting different context information, with the \emph{Epidemic} protocol measuring the number of messages that reached the destination. Since \emph{Epidemic} is a sort of controlled flooding, its successful delivery rate is very high, so our goal was to reach closer to its results.

In this experiment, a vehicle was chosen to create a message every 15 seconds, for delivery, generating a total of 30 messages. 
As seen in \autoref{grafico1} the combine \emph{DSC} (distance, speed, and category) shows an improvement of approximately 6\% when compared to the basic D (distance) and is 9\% below the performance of \emph{Epidemic}. Increasing the number of generated message to 60 with the TTL equal to 60 seconds, we get same results for \emph{D}, \emph{DS} and \emph{DSC}, but 6\% below the \emph{Epidemic} performance. Increasing the TTL to 120 seconds, the results of \emph{DSC} shows 5\% better performance than \emph{D} and \emph{DS}, but 12\% below \emph{Epidemic}. We compared both \emph{Epidemic} and \emph{DSC}, with 60 generated messages, increasing message TTL to 150 seconds, and the results displayed in \autoref{grafico2} show 100\% delivery success rate to \emph{Epidemic} and 98\% to our approach.

@@

This paper presented \emph{MFCV} - a message forwarding protocol for VANETs. \emph{MFCV} exploits context information to improve the opportunistic forwarding of messages. The vehicle distance to the destination, its speed and categories (type and use), were used in the decision making to enhance the process. Vehicle usage category was used to extract mobility habits of each vehicle. We used path history of all vehicles to observe if they are approaching destination and \emph{RateTimeToSend} mechanism to control the messages forwarding rate and the number of hops. We compared the results of our protocol with \emph{Epidemic} routing protocol, in terms of successful delivery rate, average latency and the number of message packets send per minute. \emph{MFCV} performed only 6 to 9\% worse than \emph{Epidemic} in terms of message delivery success. Additionally, in terms of average latency, our protocol performed 10 to 15\% better than \emph{Epidemic}. Finally, in terms of message packets sent, our protocol sent around 70\% fewer message packets, as compared to \emph{Epidemic}, in the worst case. The results concluded that our protocol performed as good as \emph{Epidemic}. Moreover, it consumed fewer resources as well.

As a future work, we intend to evaluate the impact of buffer size, the number of messages generated, and the number of vehicles using real-world mobility traces. in our forwarding scheme. Additionally, we intend to investigate other vehicles categories such as buses and trucks in different environments like city and highways. We also plan to exploit other context information such as acceleration and heading of vehicles to increase the success rate of message delivery and reduce latency.
